---
title: "Cardiotoxicity analysis"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: kable
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
    margin-top: 100px;
    margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px !important;
}

th, td { padding: 5px; }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library}
library(tidyverse)
library(gtsummary)
library(survival)
theme_gtsummary_compact()
```

I created a new column `doxo_dose_categories` with 2 categories:  
- if ctox_ag_cum_dose strictly > 240, the categorie is named `> 240mg/m2`  
- if ctox_ag_cum_dose <= 240, the categorie is named `≤ 240mg/m2`  
```{r}
card_data <- 
  readxl::read_xlsx(paste0(here::here(), "/Diep_100_calls__.xlsx"),
                    sheet = "Dox Subpopulation", n_max = 100) %>% 
  janitor::clean_names()

card_data <- card_data %>% 
  mutate(cardtox = case_when(
    cardtox == 1          ~ "Cardiotoxicity",
    cardtox == 3          ~ "No Cardiotoxicity"
  ), cardtox = factor(cardtox, levels = c("No Cardiotoxicity", "Cardiotoxicity"))
  ) %>% 
  mutate(vitals = case_when(
    vital == 1            ~ 1,
    vital == 2            ~ 0
  )) %>% 
  mutate(race = case_when(
    race == 1            ~ "White",
    race == 2            ~ "Others"
  )) %>% 
  mutate(gender = case_when(
    gender == 1            ~ "Male",
    gender == 2            ~ "Female"
  ), gender = factor(gender, levels = c("Male", "Female"))
  ) %>% 
  mutate(october_22_chip = case_when(
    october_22_chip == 1            ~ "CH",
    october_22_chip == 2            ~ "No CH"
  ), october_22_chip = factor(october_22_chip, levels = c("No CH", "CH"))
  ) %>% 
  mutate(c_type = case_when(
    c_type == 1                     ~ "breast",
    c_type == 2                     ~ "blood",
    c_type == 3                     ~ "ovarian",
    c_type == 4                     ~ "sarcoma",
    c_type == 5                     ~ "other"
  )) %>% 
  mutate(c_type_breast_vs_others = case_when(
    c_type == "breast"              ~ "breast",
    c_type == "blood"               ~ "others",
    c_type == "ovarian"             ~ "others",
    c_type == "sarcoma"             ~ "others",
    c_type == "other"               ~ "others",
  ), c_type_breast_vs_others = factor(c_type_breast_vs_others, levels = c("others", "breast"))) %>% 
  mutate(hormone_tx = case_when(
    hormone_tx == 1            ~ "Yes",
    hormone_tx == 2            ~ "No"
  ), hormone_tx = factor(hormone_tx, levels = c("No", "Yes"))
  ) %>% 
  mutate(radio_heart = case_when(
    radio_heart == 1            ~ "Yes",
    radio_heart == 2            ~ "No"
  ), radio_heart = factor(radio_heart, levels = c("No", "Yes"))
  ) %>% 
  
  mutate(hx_dm = case_when(
    hx_dm == 1            ~ "Yes",
    hx_dm == 2            ~ "No"
  ), hx_dm = factor(hx_dm, levels = c("No", "Yes"))
  ) %>% 
  mutate(hx_hld = case_when(
    hx_hld == 1            ~ "Yes",
    hx_hld == 2            ~ "No"
  ), hx_hld = factor(hx_hld, levels = c("No", "Yes"))
  ) %>% 
  mutate(hx_smk = case_when(
    hx_smk == 1            ~ "Yes",
    hx_smk == 3            ~ "Never"
  ), hx_smk = factor(hx_smk, levels = c("Never", "Yes"))
  ) %>% 
  mutate(doxo_dose_categories = case_when(
    ctox_ag_cum_dose > 240            ~ "> 240mg/m2",
    ctox_ag_cum_dose <= 240           ~ "≤ 240mg/m2"
  ), doxo_dose_categories = factor(doxo_dose_categories, levels = c("≤ 240mg/m2", "> 240mg/m2"))
  )

```

# Summary table
```{r}
card_data %>% 
  select(race, gender, october_22_chip,
         c_type, c_type_breast_vs_others,
         hormone_tx, radio_heart,
         cardtox,
         hx_dm, hx_hld, hx_smk, doxo_dose_categories) %>% 
  tbl_summary(type = list(c(hormone_tx, radio_heart,
                            hx_dm, hx_hld) ~ "categorical")
  )
```


# Regression
```{r}
glm(cardtox ~
      age_ctox_in + ctox_ag_cum_dose + gender + october_22_chip + hx_dm + hx_hld + hx_smk + race,
    data = card_data,
    family = "binomial"
) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE,
                 add_estimate_to_reference_rows = TRUE)
```

```{r}
card_data <- card_data %>% 
  mutate(gender = factor(gender, levels = c("Female", "Male")))

glm(cardtox ~
      age_ctox_in + ctox_ag_cum_dose + gender + october_22_chip + hx_dm + hx_hld + hx_smk + race,
    data = card_data,
    family = "binomial"
) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE,
                 add_estimate_to_reference_rows = TRUE)
```

# Survival analysis
```{r}
tbl1 <- card_data %>% select(vitals, surv_time,
                             c_type_breast_vs_others, hormone_tx, radio_heart) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = card_data$surv_time,
                             event = card_data$vitals)),
                   exponentiate = TRUE) %>% 
  bold_labels() %>% italicize_levels() %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level")
tbl2 <-
  coxph(Surv(time = card_data$surv_time,
             event = card_data$vitals) ~ 
          c_type_breast_vs_others + hormone_tx + radio_heart,
        data =  card_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05) %>% 
  add_nevent(location = "level") %>% add_n(location = "level")
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```



# Regression USING `doxo_dose_categories`
```{r}
glm(cardtox ~
      age_ctox_in + doxo_dose_categories + gender + october_22_chip + hx_dm + hx_hld + hx_smk + race,
    data = card_data,
    family = "binomial"
) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE,
                 add_estimate_to_reference_rows = TRUE)
```

# Univariable
```{r}
glm(cardtox ~
      doxo_dose_categories,
    data = card_data,
    family = "binomial"
) %>%
  tbl_regression(exponentiate = TRUE,
                 intercept = TRUE,
                 add_estimate_to_reference_rows = TRUE)
```


